Literature review of Formal Methods applied to Blockchain technology

### 2022-01-11 article/article.tex
finish formal application

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Software Quality, Formal Methods, Blockchain, Smartcontracts}

@@

\section{Formal Modeling Application}
Blockchain systems can be divided into five main layers~\cite{Duan2018}: Blockchain system, Consensus, Smartcontract and Application. 
As seen on Figure~\ref{fig:blockchain-layers}.

@@

>>>>

\usepackage{listings}

@@

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Software Quality, Formal Methods, Blockchain, Smart contracts}

@@

\section{Formal Modeling Application}
Blockchain systems can be divided into five main layers~\cite{Duan2018}: Data layer, Consensus layer, Smart contract layer and Application layer. 
As seen on Figure~\ref{fig:blockchain-layers}.

@@

The data layer is responsible for block generation, the construction of the blockchain and storage~\cite{Duan2018}. 
The network layer connects the data and filters packets, and also manages the nodes. 
The rules of how the blockchain protocol must work are defined on the consensus layer. 
This includes agreement between nodes, fault tolerance and data consistency. 
Smart contracts have their execution lifecycle on the smart contact layer.
The application layer is the upper level layer of used applications such as games or decentralized financial services.

It is possible to use languages to define the standards of a blockchain system with mathematical rigor.
One of them is SDL (Specification and Description Language), an Object-Oriented formal language defined by the International Telecommunications Unit.
It is capable of mathematically describing the structure, behavior, and data of real-time and distributed communicating systems. 
It is also used by IEEE for standard definition~\cite{Duan2018}.

With SDL, we can describe a structure at the levels of system, block or process. A system represents the object to describe. 
It is responsible for the communication with the outer environment through the appropriate channels.
A system can contain one or more blocks and blocks contain one or more processes.
The majority of the functionality of an SDL system is described on the process level.
Comparing to conventional programming languages, modules are represented by blocks in SDL, and funcionalities are represented by processes.
In~\cite{Duan2018}, the authors propose a hierarchical model focused on the consensus layer.

  \subsection{Data Structure}
  In~\cite{Duan2018} the modelled blockchain system used an Improved Byzantine Fault Tolerant Algorithm as the consensus algorithm.
  To define the data structure of a block in SDL, we need to represent them as a pair with a header BT and body BH, where:

  \begin{displaymath}
    B~\equiv(B_H,B_T) \quad \textrm{with} \quad B_H~\equiv(\mathbb{H}_p,\mathbb{H}_a,\mathbb{H}_t,\mathbb{H}_m,\mathbb{H}_h) 
  \end{displaymath}
  \begin{displaymath}
    \quad \textrm{and} \quad B_T~\equiv(T_1,T_2,...,T_{\mathbb{H}_l}) 
  \end{displaymath}
  \linebreak[1]

  In these expressions, $\mathbb{H}_p$ represents the hash value of the previous block, 
  $\mathbb{H}_p$ the hash value of the current block, $\mathbb{H}_t$ it's timestamp, 
  $\mathbb{H}_m$ the merkle tree root (the hash of the transactions of a block), and $\mathbb{H}_h$ the current block's height.

  The the block' structure definition can be seen on Listing~\ref{lst:block-code}.

  \begin{lstlisting}[frame=single,label={lst:block-code},
    caption={Definition of a block's structure},captionpos=b]
    newtype blockchain struct
      prehash integer;
      hash integer;
      length integer;/*the length of blocks*/
      merkleroot integer;
      ti Time;/*Timestamp*/
      translist list;
    endnewtype;

    newtype list
     array(maxit, integer)
    endnewtype;

    syntype maxit =Integer constants 0:25
    endsyntype;
  \end{lstlisting}

  A blockchain can be regarded as a state machine~\cite{Duan2018} that contains a starting state, a non-empty state, the input transaction set,
  a state transition function, and an acceptance state set. Which can be formally defined as:
    \begin{displaymath}
      S~\equiv(Q,\Sigma,\delta,s,F)
    \end{displaymath}

  Where:

  \begin{enumerate}
    \item[$\blacksquare$] Q is the non-empty state set, which represents all the states.
    \item[$\blacksquare$] $\Sigma$ is the set of the newly generated and consensus blocks. 
    \item[$\blacksquare$] $\delta$ represents the state transition function,
      \begin{math}
        \delta: Q \times \Sigma \rightarrow Q
      \end{math}.
    \item[$\blacksquare$] s would be the starting state, which is the state of the system when it initializes.
      \begin{math}
        \delta: s \in Q
      \end{math}.
    \item[$\blacksquare$] F representes the acceptance state set.
      \begin{math}
        F \subseteq Q
      \end{math}.
  \end{enumerate}
  \  \\ %fake break line
  The state of the model at the first block generation is the starting state.
  As transactions are made, the model a leader node to generate the block, which is transmitted to the other nodes on the blockchain network,
  triggering its validation for consensus. 
  A consensus block stores the hash of the previous one and adds it to the end of the blockchain and completes the transfer of the blockchain state.

  \subsubsection{Diagrams}
  With a modelling language like SDL, we can use diagrams to define blockchain structural entities, like the core system of the consensus layer
  and blocks. Process flowcharts can model the flow of information and input validation, making a process like
  node validations of blocks and its subprocess of synchronization and voting can clearly illustrated.

\section{Consensus Rule}
Blockchain system may choose to use consensus rules taken from popular protocols such as Bitcoin,
but an alternative rule can be implemented, and also be formally modelled~\cite{Kawahara2020}.

On a consensus rule, fault tolerances requirements are of utmost importance, including the collusion of nodes to alter a ledger.
In~\cite{Kawahara2020}, the author discusses a way to describe these tolerances with a scalable verification.
